Lyman-$\alpha$ photons through rotating outflows

### 2018-03-31 paper/paper.tex
edit

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
kinematics of galaxies.
Here we perform Monte Carlo radiative transfer simulations of outflowing
gas with additional solid body rotation to understand how these kinematic
features impact the morphology of the Lyman-$\alpha$ emission line.
We explore a range of Hydrogen optical depth of
$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq 100$ and
outflow velocity $0\leq \vout/\kms\leq 50$.  
We find three important consequences of rotation.
First, it introduces a dependency with viewing angle; second it
produces a line broadening and third it increases the flux at the
line's center.
We also develp a semi-analytic model that modifies the spectra of
pure outflow simulations that manages to reproduce the line widths
and flux change at the line's center within a $7\%$ and $50\%$
precision, respectively.
\end{abstract}

@@

The interpretation of the Lyman-$\alpha$ emission in galaxies is
key to understand their evolutionary processes, specially in
star-forming, low-dust galaxies \citep{PartridgePeebles}.
Recent improvements in instrumentation have revolutionized the kind of
studies that can be performed on Lyman-$\alpha$ emitting galaxies
(LAEs.)
For instance, it is now possible to infer detailed kinematic maps for
nearby galaxies.
The study of these maps would allow us to build data-driven models to
interpret the \lya spectra of unresolved galaxies, helping us to
constrain the physical conditions of the interstellar medium (ISM)
processing the \lya radiation.

For instances, dwarf galaxies show signs of coherent rotation
\citep{2009A&A...493..871S}.
Some of them are expected to have high neutral gas contents and thus
show \lya\ emission with rotation imprints
\citep{2005A&A...433L...1B,2008ApJ...672..888T,2013MNRAS.434.2491G}
Some Compact Dwarf Galaxies actually show gas kinematics that range
from pure rotation to high velocity dispersion without a clear
rotation pattern.
\citep{2015A&A...577A..21C,2017A&A...600A.125C}

The case for outflows as traced in the \lya\ line is much better
stablished.
In many cases the \lya line profiles has a single peak
redwards from the line's center, in other cases there is a double peak
but the peak on the red side is stronger
\citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.   
These features been readily explained as the consequence of multiple
\lya photon scatterings through an homogeneous outflowing shell of
neutral Hydrogen
\citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.

Here we present for the first time a study of the joint effects of
galaxy outflows and rotation, two of the three major kinematic
components expected in LAEs.
We study a simplified geometrical configuration corresponding to a
spherical gas cloud with symmetrical radial outflows and a rotation
profile corresponding to a solid body.
We base our modeling on a Monte-Carlo radiative transfer code called
CLARA (Code for Lyman Alpha Radiation Analysis) presented for the
first time by \cite{CLARA}.

Besides modeling the impact of joint rotation and outflows, we also
check to what extent the analytical model presented by
\cite{Garavito14} to explain the effects of rotation can also be
applied in our case.
We show how solid body rotation effects can be modelled by
postprocessing the results of outflow only simulation.

in the results of other radiative transfer model in the case of large
optical depths.
In this case it is a good approximation to Doppler
boost the results of the model without rotation.

>>>>

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
kinematics of galaxies.
Here we perform Monte Carlo radiative transfer simulations of outflowing
gas with additional solid body rotation to understand how these kinematic
features impact the morphology of the Lyman-$\alpha$ emission line.
We explore a range of Hydrogen optical depth of
$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq 100$ and
outflow velocity $0\leq \vout/\kms\leq 50$.  
We find three important consequences of rotation.
First, it introduces a dependency with viewing angle; second it
produces a line broadening and third it increases the flux at the
line's center.
We also develp a semi-analytic model that modifies the spectra of
pure outflow simulations and manages to reproduce the line widths
and flux change at the line's center within a $7\%$ and $50\%$
precision, respectively.
\end{abstract}

@@

Recent improvements in instrumentation have revolutionized the kind of
kinematic studies that can be performed on galaxies with small angular
sizes on the sky.
For instance, spatially resolved spectra on dwarf galaxies show signs
of coherent rotation \citep{2009A&A...493..871S}.
Some Compact Dwarf Galaxies actually show gas kinematics that range
from pure rotation to high velocity dispersion without a clear
rotation pattern.
\citep{2015A&A...577A..21C,2017A&A...600A.125C}
These studies are now being extended to distant galaxies.

Some of those galaxies are expected to have high neutral gas contents
and thus show \lya\ emission with rotation imprints
\citep{2005A&A...433L...1B,2008ApJ...672..888T,2013MNRAS.434.2491G}
The computational modeling of systems showing rotation has only been
studied until recently by \cite{Garavito14}.
In that work the authors studied the influence of pure solid body
rotation on the \lya line's morpholgy.
They found that rotation indeed introduces changes, the most
noticeable change is the dependence of the spectra with the viewing
angle with respect to the rotation axis.

\cite{Garavito14} also presented a simple semi-analytical
approximation that accounted for the main features of the \lya spectra
from a rotation sphere.
Recently, this semi-analytic solution was used to perform a Markov
Chain Monte Carlo exploration to fit the observed spectra Compact
Dwarf Galaxy \citep{tololo}, that presented atypical features that
could be explained by pure rotation.

However, the gas dynamics in Lyman Alpha Emitter (LAE) galaxies is
more complex than pure rotation.
The case for outflows as traced in the \lya\ line is much better
stablished.
In many observations the \lya line profiles has a single peak
redwards from the line's center, in other cases there is a double peak
but the peak on the red side is stronger
\citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.   
These features been readily explained as the consequence of multiple
\lya photon scatterings through an homogeneous outflowing shell of
neutral Hydrogen
\citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.

An study of the combined effects of outflows and rotation has not been
presented in the literature.
Here we present for the first time such as study. 
We study a simplified geometrical configuration corresponding to a
spherical gas cloud with symmetrical radial outflows and solid body
rotation. 
We base our modeling on a Monte-Carlo radiative transfer code.
Besides modeling the impact of joint rotation and outflows, we also
check to what extent the analytical model presented by
\cite{Garavito14} to explain the effects of rotation can also be
applied in our case.
